Clicks and Cliques

### 2018-12-13 papersoul.Rnw
update data chunk, map is not working

# packages
library(reshape2)
library(plyr)
library(productplots) 
library(ggplot2)
library(stats)
library(randomForest)
library(dplyr)
library(tidyr)
library(maps)
library(fpc)
library(xtable)
library(car)
library(GGally)
library(survey)
library(reldist)

@@

<<data,echo=FALSE,warning=FALSE,message=FALSE>>=
# Procesing data sets from SOCT and Census Bureau
source('code/data_chk.R')
source('code/census_chk.R')

@@

<<data2, cache=TRUE,echo=FALSE,warning=FALSE,message=FALSE,eval=FALSE>>=
# Data contain on 'sotc.csv' file were pre-processed by Xiaoque Cheng with mergeGUI.
soul <- read.csv('data/sotc.csv', header=T)

# Variable sets
orig <- c(4:6, 8:61 ,148:152, 165:175, 181:206, 210:213)
demo <- c(7, 62:79, 153:154, 158:160, 206:209, 216, 229)
reco <- c(80:120, 141:147, 161, 177:179, 217:228)
cons <- c(121:135, 155:157, 162:164, 176, 180)

# recode CCEGRP
soul$CCEGRP2 <- recode(soul$CCEGRP,"'Engaged'='Attached';'Not Engaged'='Not Attached'")

# Compute attachment proportion for ordering QSB levels
wt     <- with(subset(soul,!is.na(WEIGHT)), WEIGHT/3)
des.wt <- svydesign(ids=~1, strata = ~QSB, weights = ~wt, data = subset(soul, !is.na(WEIGHT) ) )
att.pr <- svyby(~CCEGRP2, ~QSB, des.wt, svymean, keep.var = TRUE, na.rm = T)
soul$QSB2 <- soul$QSB
soul <- merge(soul,att.pr[,1:2], by = 'QSB')
soul$QSB <- with(soul, reorder(QSB2, CCEGRP2Attached)  )
soul$CCEGRP2Attached <- NULL

# Reduction of the data set to avoid missing value imputation.
miss <- is.na(soul)
aux  <- apply(miss, 2, mean )

# select variables with less than 25 % of missing values
var.red <- names(aux[aux < .25])
aux     <- apply(miss[,var.red], 1, sum )
aux1    <- data.frame(table(aux))
soul.red <- soul[aux < 3, var.red]
orig.red <- colnames(soul.red)[c(4, 5, 7, 8:50)]

# age and proportion of age in place where live.
soul <- mutate(soul, QD1gr = cut(QD1, breaks = c(0, 30, 40, 50, 60, 70, 80, 99), include.lowest = T),
               QD2gr=cut(QD2, breaks = c(0, 10, 20, 50, 99), include.lowest = T), prop.res = QD2/QD1,
               prop.gr=cut(prop.res, breaks = c(0, 0.25, 0.5, 1, 5.5), include.lowest = T))

soul$prop.res[soul$QD1 == 0] <- NA

demo.red <- c(colnames(soul.red)[c(6, 51:62)],'QD1gr','QD2gr','prop.res','prop.gr')

# Main reduction of data: Only variables with less than 25% of missing, and complete cases withion those vars.
aux <- apply(is.na(soul[,c(orig.red,demo.red,'CCE','CCEGRP2')]), 1, sum)
condi1 <- aux==0
soul.red <- soul[condi1,c(orig.red,demo.red,'CCE','CCEGRP2','QSB','YEAR','WEIGHT','PROJWT')]

# Recodification of several variables in the data, collapsing some levels, and considering NA some of the responses (like, DK, not aplicable, non-response, etc).
reco.fun <- function(x) {
  n <- length(x)
  y <- numeric(n)
  if ("2" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)","(Have not lived in area for five years)")] <- NA
    y[x %in% c("Very bad","Strongly disagree","Extremely low","Not at all safe","Not at all satisfied","Not at all likely","Much worse","Will be much worse")] <- 1
    y[x %in% c("Very good","Strongly agree","Extremely high","Completely safe","Extremely satisfied","Extremely likely","Much better","Will be much better")] <- 5
    y[x=='2'] <- 2 ; y[x=='3'] <- 3 ; y[x=='4'] <- 4
  }
  if ("Yes" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)")] <- NA
    y[x=='Yes'] <- 1 ; y[x=='No'] <- 0
  }
  if ("A few" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)")] <- NA
    y[x =='None'] <- 1; y[x =='A few'] <- 2 ; y[x =='Some'] <- 3;y[x =='About half'] <- 4;
    y[x %in% c("Most","Most, OR")] <- 5; y[x =='All or nearly all'] <- 6
  }
  if ("Never" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)")] <- NA
    y[x %in% c('Never', 'NeverNunca')] <- 1; y[x =='Once a year or less'] <- 2
    y[x %in% c('Once a month','Several times a year')] <- 3;y[x =='Once a month'] <- 4;
    y[x %in% c('Several times a week','About every day')] <- 6; y[x =='Several times a month'] <- 5
  }
  if ("Decreased" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)")] <- NA
    y[x =='Decreased'] <- 1; y[x =='Stayed the same'] <- 2;  y[x =='Increased'] <- 3
  }
  if ("Getting better" %in% levels(x) ) {
    y[x %in% c("(DK)","(Refused)","(Not applicable)")] <- NA
    y[x =='Getting worse'] <- 1; y[x =='Getting better'] <- 3;  y[x =='(Same)'] <- 2
  }
  if (is.numeric(x) ) y <- x
  return(y)
}

reco.aux <- NULL
for (i in 1:length(orig.red)) reco.aux <- rbind(reco.aux, reco.fun(soul.red[,i]) )
reco.aux <- t(reco.aux)[,-32]
colnames(reco.aux) <- orig.red[-32]

# More recodification, now for demographic variables in the data.
reco.demo <- soul.red[,c(demo.red,'Q11')]

reco.demo <- mutate(reco.demo,
              Q11=factor(recode(Q11,"'(Disabled/unable to work)'='Other ';'(Other) (do not list)'='Other '
                                                ;'Disabled/unable to work'='Other ';'Temporarily laid off'='Other '
                                                ;'Unemployed and not looking for work'='Other '; '(DK)'='Non Response';'
                                                (Refused)'='Non Response' ")),
              QD4=factor(recode(QD4,"'(DK)' = 'Non Response' ;'(Refused)'='Non Response'")),
              QD6=factor(recode(QD6, "c('Separated','Widowed','Separated, OR')='Divorced';'Never been married'
                                       ='Single'; 'Living in a partnered relationship' = 'Married'; 'Now married'='Married'
                                       ;'(DK)'='Non Response';'(Refused)'='Non Response'")),
              QD7=factor(recode(QD7, "'Grade school or less'='Less HS'; 'Some high school'='Less HS';'High school
                                      graduate'='HS';'(DK)'='Non Response';'(Refused)'='Non Response';c('Some college or
                                      technical school','College graduate','College graduate, OR')='Collegue';'Post
                                      -graduate work or degree'='MS'")),
              QD8=factor(recode(QD8, "'Other (rent a room, live as a lodger, squatter, etc.)'='Other';'(DK)'='Non
                                      Response';'(Refused)'='Non Response'") ),
              QD9=factor(recode(QD9, "c('Under $15,000','$15,000 to $24,999','$25,000 to $34,999') = 'Under 35k'; c
                                      ('$35,000 to $44,999','$45,000 to $54,999','$55,000 to $74,999')='35k to 75k'; c('$75
                                      ,000 to $99,999','$100,000 or over')='Over 75k'")))

# QD10 ask for hispanic or not hispanic, QD111 ask for race, we combine both in QD111
aux <- reco.demo$QD111
levels(aux)[c(4, 6:10, 15:17, 19, 20) ] <- 'Other'
levels(aux)[c(1, 2, 5, 7, 8, 10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[reco.demo$QD10=='Yes'] <- 'Hispanic'
reco.demo$QD111  <- factor(aux)

# new soul.red data with recode variables and no missing
soul.red2 <- data.frame(reco.aux, reco.demo,soul.red[,c('CCE','YEAR','CCEGRP2','QSB','WEIGHT','PROJWT')])
soul.red2$Q23[soul.red2$Q23==9] <- NA

# the recodification creates a few new missing values that we don't want.
aux <- apply(is.na(soul.red2), 1, sum)
condi2 <- aux==0
soul.red2 <- soul.red2[condi2 , ]
@

@@

soul.red2 <- mutate(soul.red2, CCE = soul$CCE[condi1][condi2],
                    URBAN_GR = soul$URBAN_GR[condi1][condi2],
                    QSB = soul$QSB[condi1][condi2],
                    WEIGHT = soul$WEIGHT[condi1][condi2])

wt <- with(subset(soul.red2,prop.res<=1), WEIGHT/3)
des.wt <- svydesign(ids=~1, strata=~QSB, weights=~wt, data=subset(soul.red2,prop.res<=1))

life <- svyby(~prop.res+QD1+CCE, ~QSB, des.wt, svymean, keep.var=TRUE, na.rm=T)
lifevar <- svyby(~prop.res+CCE, ~QSB, des.wt, svyvar, keep.var=TRUE, na.rm=T)
lifevar$rho <- with(lifevar, V2/sqrt(V1*V4))
rho <- with(life, round(cor(CCE, prop.res),2) )

qplot(data=life, x=prop.res, y=QSB, size=I(4), ylab='',xlab='Proportion of lifetime living in the community')+theme(axis.text.y = element_text( size=I(10)))

@
\caption{Dot chart of communities sorted from top to bottom by attachment against the mean proportion of lifetime spent in the current community (stay). Communities where most of people are living longer are the less attached ones. \label{stay}}
\end{figure}

<<life.tab, dependson=c('corplot','life'),echo=FALSE, results='asis'>>=
aux1 <- subset(dat.cor, vars %in% c("prop.new.entry","av.rent","av.own" ) & type=='Attached')[,1:2]
aux2   <- data.frame(vars='stay', cor=rho)
aux3 <- rbind(aux1,aux2)
description <-c('Average years renting in current home', 'Average years owning the current home',
              'Prop. of people not borned in the community', 'Proportion of age in current comunity')
data.source <- c(rep('Census Bureau', 3), 'Knight-Gallup data')
tab <- data.frame(Data=data.source,Description=description,Correlation=aux3$cor)
row.names(tab) <- NULL
tab <- xtable(tab, caption='Correlation with attachment', label='lifetab')
print(tab, caption.placement='top')
@

@@

<<moving, dependson='data', echo=FALSE,results='asis'>>=

levels(soul$Q5)[1:2] <- 'Non Response'
levels(soul$Q5)[2:6] <- 'Move from your neighborhood'
soul$Q5 <- factor(soul$Q5)

@@

\begin{figure}
<<life_seq, dependson=c('life'), echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=

life$QSB <- with(life, factor(QSB, levels = rev(levels(QSB))))

qplot(data=lifevar,x=QSB,ymax=rho,ymin=0, xlab='',geom='linerange', ylab='Correlation',size=I(2))  + theme(axis.text.y = element_text( size=I(10) ),axis.text.x = element_text(size=I(10), angle=90),axis.title.x = element_text(size=I(25)))

@
\caption{ Bar chart of correlations between attachment and lifetime living in the community within each community (ordered by attachment level). Only few communities shows negative correlation between attachment and lifetime living there. \label{wrho}}
\end{figure}

@@

# QD10 ask for hispanic or not hispanic, QD111 ask for race, we combine both in QD111
aux <- soul$QD111
levels(aux)[c(4,6:10,15:17,19,20) ] <- 'Other'
levels(aux)[c(1,2,5,7,8,10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[reco.demo$QD10=='Yes'] <- 'Hispanic'
soul$race  <- factor(aux)

>>>>

# packages

# to be removed...
library(reshape2)
library(plyr)

# data wrangling
library(tidyverse)
library(stats)
library(car)

# stat vizualization
library(productplots) 
library(ggplot2)
library(GGally)
library(maps)

# stat methods
library(randomForest)
library(survey)
library(fpc)
library(reldist)

# write report
library(xtable)

@@

<<data,echo=FALSE,warning=FALSE,message=FALSE, cache=FALSE>>=
# Procesing data sets from SOCT and Census Bureau
source('code/data_chk_tidy.R')
source('code/census_chk.R')

@@

@@

### VARIFICAR NOMBRES REPETIDOS EN SVYVAR !!!!!!!
life <- svyby(~prop.res+QD1+CCE, ~QSB, des.wt, svymean, keep.var=TRUE, na.rm=T)
lifevar <- svyby(~prop.res+CCE, ~QSB, des.wt, svyvar, keep.var=TRUE, na.rm=T)
lifevar$rho <- with(lifevar, V2/sqrt(V1*V4))

qplot(data=life, x=prop.res, y=QSB, size=I(4), ylab='',xlab='Proportion of lifetime living in the community')+theme(axis.text.y = element_text( size=I(10)))
@
\caption{Dot chart of communities sorted from top to bottom by attachment against the mean proportion of lifetime spent in the current community (stay). Communities where most of people are living longer are the less attached ones. \label{stay}}
\end{figure}

<<life.tab, dependson=c('corplot','life'),echo=FALSE, results='asis', cache=FALSE>>=
aux1 <- subset(dat.cor, vars %in% c("prop.new.entry","av.rent","av.own" ) & type=='Attached')[,1:2]

rho <- with(life, round(cor(CCE, prop.res),2) )

aux2   <- data.frame(vars='stay', cor=rho)
aux3 <- rbind(aux1,aux2)
description <-c('Average years renting in current home', 'Average years owning the current home',
              'Prop. of people not borned in the community', 'Proportion of age in current comunity')
data.source <- c(rep('Census Bureau', 3), 'Knight-Gallup data')
tab <- data.frame(Data=data.source,Description=description,Correlation=aux3$cor)
row.names(tab) <- NULL
tab <- xtable(tab, caption='Correlation with attachment', label='lifetab')
print(tab, caption.placement='top')
@

@@

<<moving, dependson='data', echo=FALSE,results='asis', cache=FALSE>>=
levels(soul$Q5)[1:2] <- 'Non Response'
levels(soul$Q5)[2:6] <- 'Move from your neighborhood'
soul$Q5 <- factor(soul$Q5)

@@

\begin{figure}
<<life_seq, dependson=c('life'), echo=FALSE,fig.height=4,message=FALSE,warning=FALSE>>=
life$QSB <- with(life, factor(QSB, levels = rev(levels(QSB))))

lifevar %>% data.frame() %>%  select(-starts_with('se')) %>%
  qplot(data=., x=QSB,ymax=rho,ymin=0, xlab='',geom='linerange', ylab='Correlation',size=I(2))  + theme(axis.text.y = element_text( size=I(10) ),axis.text.x = element_text(size=I(10), angle=90),axis.title.x = element_text(size=I(25)))
@
\caption{ Bar chart of correlations between attachment and lifetime living in the community within each community (ordered by attachment level). Only few communities shows negative correlation between attachment and lifetime living there. \label{wrho}}
\end{figure}

@@

# QD10 ask for hispanic or not hispanic, QD111 ask for race, we combine both in QD111

aux <- soul$QD111
levels(aux)[c(4,6:10,15:17,19,20) ] <- 'Other'
levels(aux)[c(1,2,5,7,8,10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[soul$QD10=='Yes'] <- 'Hispanic'
soul$race  <- factor(aux)
